%*************************************************************%
%
%    Ensemble, 1_42
%    Copyright 2003 Cornell University, Hebrew University
%           IBM Israel Science and Technology
%    All rights reserved.
%
%    See ensemble/doc/license.txt for further information.
%
%*************************************************************%
\section{Configuration}

%----------------------------------------------------------------------

\subsection{Command-line Arguments and Environment Variables}
\label{subsect:config}
\ensemble\ applications typically support a variety of configuration
parameters.  Most of these can be configured through command-line options as
well as through setting environment variables.  In all cases, command-line
options override environment variables.  Look in \sourceappl{arg.ml} for the
authoritative list of the configuration parameters.  Some are listed below as
command-line options.  The corresponding configuration variable for
\mlval{-group\_name} (for example) is \mlval{ENS\_GROUP\_NAME} (the name is
capitalized and the `-' is replace with \mlval{ENS\_}).
\begin{description}
\item
[\mlval{-modes} arg :] Set the default modes for an application to use.  The
modes are specified giving their names in all-uppercase, each separated by
single colons (`:') and no white-space.
\item
[\mlval{-udp\_port} port :] set the default UDP port for \ensemble\
applications.  For point to point UDP communication, this is the port
number \ensemble\ first tries to bind to for UDP communication (if it
is already in use Ensemble will then fail).  It can be set to any
value.  The default is to let the operating system choose a port to
use.
\item
[\mlval{-host\_ip} IP address:] set the default IP address of the host for \ensemble\
applications. This allows overriding the default IP-address of the
machine, a useful option if the machine is connected to several
network interface cards (NICs) and has several IP addresses. 
\item
[\mlval{-deering\_port} port :] This is the port that \ensemble\ will
use for Deering IP multicast communication (if enabled).  All
processes must use the same port number.
\item
[\mlval{-gossip\_port} port :] sets the port that the \mlval{gossip} servers
use.
\item
[\mlval{-gossip\_hosts} arg :] sets the hosts where applications using UDP
communication can look for \mlval{gossip} servers.  The value should be a
colon-separated list of hostnames.  The \mlval{gossip} server application will
only execute on these hosts.  Note that you only have to execute a gossip
server on one of these hosts: applications will try each of the hosts in turn
while looking for a gossip server.  However, multiple servers can be executed
for increased availability.
\item
[\mlval{-id} name :] used to give applications unique identifiers.
Usually this is set to be your user id.  Setting this variable
prevents \ensemble\ applications run by other users from interacting
with yours.  In case you do want them to interact, you should set
their variables to have the same value.  If using DEERING IP
multicast, their \mlval{-deering\_port} variable should also be set to
the same value.
\item
[\mlval{-groupd\_port} port :] sets the port that the membership \mlval{groupd}
servers use.
\item
[\mlval{-groupd\_hosts} arg :] sets the hosts that the membership
\mlval{groupd} servers use.  Format is the same as for \mlval{-gossip\_hosts}.
\item
[\mlval{-groupd} :] Use the membership service on the local host (see
section~\ref{section:groupd}.  This option may override others.
\item
[\mlval{-group\_name} name :] Set the name of the application's group.
\note{Currently, only the \mlval{ensemble} application supports this.}
\item
[\mlval{-key} key :] Set the key to use for a particular application.  All
messages sent and received by the application will be authenticated with this
key.
\item
[\mlval{-secure} :] Enable security enforcement.  This prevents any insecure
communication transports from being initialized.
\item
[\mlval{-add\_prop} property :] Adds a specific property to the
Ensemble protocol stack. See Section~\ref{sec:properties} for more
information on supported Ensemble properties. 
\item 
[\mlval{-remove\_prop} property :] The dual of add\_prop.
\item 
[\mlval{-sock\_buf} size :] The size of socket buffers to request
from the operating system.  The default size is $52428$ (the
traditional limit on Unix).  If you are using \ensemble\ in
high-performance setting and are experiencing message loss, this is a
parameter that should be increased.
\item
[\mlval{-multiread} :] Enable multiple reads on sockets.  The default
is to receive and process one message from the operating system
at a time.  Setting this will cause all available messages to be read
from sockets before processing any of them, which may reduce message
loss due to buffer overflow in the operating system.
\item
[\mlval{-pollcount} count :] The number of times to query the
operating system before blocking.  \ensemble\ blocks after checking
(via the \mlval{select()} system call) the operating system for
messages and not finding any.  Setting this to $1$ will cause
\ensemble\ to block immediately when there are no more messages.
Setting this to a large number will cause \ensemble\ to busy-poll for
a longer time before blocking.
\item
[\mlval{-ranking} trans:] this is used to set the priority of
transport mechanisms. This is useful if you wish to use TCP as a
transport instead of UDP (the default). Using {\tt -ranking TCP} will
cause TCP to be ranked higher than any other transport. 
[\mlval{-trace} module\_name:] Trace a specific module, it is possible
to trace several modules. Module naming is insensitive to lower/upper casing.
\end{description}

The following configuration parameter can only be set as an
environment variable.
\begin{description}
\item
[\mlval{ENS\_TRACE}]: enables module initialization tracing.  With this set (to
any value), modules print out their names as they initialize.  This is useful
if an exception  occurs during initialization because because it
enables you to narrow the problem down to one module.
%\item
%[\mlval{ENS\_LOG\_HOST} and \mlval{ENS\_LOG\_PORT}]: sets the port and host of the remote
%log server.  Applications can execute as the server via the \mlval{-log\_server}
%command-line option.  Running programs with the \mlval{-log} option enables the
%remote logging facility for other programs.  \note{This is currently not documented:
%you have to read the source code in \sourceappl{log} in order to use it now.}
\end{description}

Prior to version 1.42 all configuration variables were defined in the
environment. This meant that the environment-variable system was
tainted with the various Ensemble configuration options. Starting with
version 1.42 configuration is handled using a configuration-file which is
by default {\tt \$USER/ensemble.conf}. To set this to a
different file use the {\tt ENS\_CONFIG\_FILE} environment variable. 

Each line in the configuration file is either (1) a comment starting with \# (2) an
empty line (3) a value binding where $ensemble\_variable=value$. For
example: 

\vspace{1mm}
\hrule{}
\begin{verbatim}
# Ensemble ID
ENS_ID=orodeh

# The port used by Ensemble for transport
ENS_PORT=6790
\end{verbatim}
\hrule{}
\vspace{1mm}

The first operation an Ensemble process performs is to open and parse
the configuration file. Note that an option set in the configuration
file takes effect for all your Ensemble processes.

\subsection{Gossip service configuration}
Not all platforms support IP-multicast. Furthermore, if you compile
Ensemble without the default native socket library you'll be using
OCaml's networking library which, up to version 3.06, did not 
support multicast. If you fall under any of these categories you'll need to
use the gossip service to allow applications to locate each other.

\mlval{ENS\_GOSSIP\_PORT} must be set to a port number that
is not used by other applications.  Normally, user applications
cannot use port numbers below 1000.  \mlval{ENS\_GOSSIP\_HOSTS} must
be set to a list of colon-separated host names where the gossip
server may be found.  If you wish to use port 7500 for the gossip
server on hosts ``ely'' and ``natasha,'' you would set these
configuration variables as follows (in the configuration file):
\begin{verbatim}
ENS_GOSSIP_PORT=7500
ENS_GOSSIP_HOSTS=ely:natasha
\end{verbatim}

Applications will require a gossip server process to be running
in order to contact each other. Before executing an application, a
gossip server must be started on one of the hosts listed in
\mlval{ENS\_GOSSIP\_HOSTS}:
\begin{verbatim}
% gossip
 ...
\end{verbatim}
On the same or other hosts execute several instances of an
application, such as \sourcedemo{mtalk}:
\begin{verbatim}
% mtalk
 ...
\end{verbatim}

\subsection{Transports}

\note{If you are only using regular IP-multicast sockets for communication,
then you do not need to read this section.}

Perhaps the most confusing part of running \ensemble\ applications
comes from selecting communication transports.  Communication
transports are the bottom-most part of \ensemble\ and are used for
sending and receiving messages on a network.  There are several ways
this can be confusing and often \ensemble\ cannot detect that there is
a problem, so you do not get a warning.  For instance, if you
configure an application so that one process is using UDP sockets for
communication and another is using NETSIM, then the two processes will
stall waiting for other processes to communicate with them on their
selected medium.

A confusing aspect of transport is that an application typically uses
two different kinds of transports: a primary transport and a gossip
transport.  Normal application communication is all done over the
primary transport, which must support point-to-point communication and
may also support multicast communication.  Communication between
different partitions of a group of applications uses the gossip
transport which must support ``anonymous'' multicast communication.

Applications occasionally send ``gossip'' messages with their gossip
transport to the rest of the ``world'' in order to inform other
partitions about their presence.  When two partitions learn of each
other, they can then merge the partitions together.  After they have
merged together, they communicate over their primary transport.  This
gossip-and-merge mechanism is used when applications first start up:
an application creates its own singleton group and then merges with
any other already existing partitions through gossiping and merging.
Thus, if there is a problem with the gossip transport, you will tend
to have a bunch of applications in singleton groups that never merge.
If there is a problem with the primary transport, the merging will
occur, but then the various members will be unable to communicate.
This will cause them to repeatedly break into partitions (when they
decide that the other members must have failed) and then re-merge
again.

The various primary and gossip transports are presented in the
following table.  The ``P'' and ``G'' columns specify whether a
transport can be used for primary communication and/or gossip
communication.
\begin{center}
\begin{tabular}{|l|c|c|l|}					   \hline
transport	& P	& G	& description			\\ \hline \hline
\mlval{UDP}	& \chk	& \chk	& UDP (+ gossip server)		\\ \hline
\mlval{DEERING}	& \chk	& \chk	& UDP/IP multicast		\\ \hline
\mlval{TCP}	& \chk	& 	& TCP/IP			\\ \hline
\mlval{NETSIM}	& \chk	& \chk	& network simulator		\\ \hline
\end{tabular}
\end{center}

The \mlval{NETSIM} transports are used only in applications that are
simulating the behavior of a group inside a single process.  The
\mlval{rand} and \mlval{fifo} demos use this, for instance. All other
currently supported modes run over IP. UDP requires running the gossip
server.  With TCP as a transport, a fully connected mesh of TCP
connections is used to move messages between group members. This is
not very efficient for multicast messages, however, it can
sometimes be useful. 

There are several ways to change the communication transports that \ensemble\
uses.  These are listed below in order of highest ``precedence.''
\begin{enumerate}
\item
Command-line argument: with the \mlval{-modes} argument (see the
command-line argument documentation).
\item
Application setting: a particular application may differ from the
\ensemble\ defaults.
\item
Environment variable: \mlval{ENS\_MODES} variable (see the environment
variable documentation below).
\item
\ensemble\ defaults: \mlval{DEERING}.
\end{enumerate}

\subsection{Using Deering IP Multicast}
Deering IP Multicast is the default configuration. If your machines support IP
multicast communication, it is preferable to use \mlval{DEERING}
transports because you will then not have to run the gossip server
with \ensemble\ applications. IP multicast is only available
when using the Socket library. It is not currently supported
by the native OCaml library. The configuration parameters pertaining
to Deering are ENS\_MODES and ENS\_DEERING\_PORT. To modify their
default settings use:
\begin{verbatim}
ENS_DEERING_PORT=1234
ENS_MODES=DEERING
\end{verbatim}
 
You can try out the IP multicast transport by using the command-line arguments.  
For example: 
\begin{verbatim}
% mtalk -modes DEERING
\end{verbatim}

\subsection{Notes and Problems}

See also the problems mentioned in the \ensemble\ reference manual.
\begin{description}
\item
[IP Multicast problems :] Some problems may occur with IP Multicast.
The time-to-live value for multicast messages may be too small in
some environments, preventing multicast messages from reaching all
members. The TTL value can be adjusted by editing the file \newline
\sourcecode{socket/s/PLATFORM/multicasts.c}.
\item
[Variation in site configuraiton :]
IP-multicast may be misconfigured or disabled for security reasons on
your site. You'll need to use the gossip service in this case. 
\item
[Multiple transports:]
Note that you can configure ENS\_MODES with serveral transports and let
Ensemble choose the best one. For example: 
\begin{verbatim}
ENS_MODES=UDP:TCP
\end{verbatim}
will cause Ensemble to use UDP or TCP, whichever is appropriate. The
system ranks Deering the best transport, then UDP, and finally TCP. In
order to modify this use the -ranking flag. For example: 
\begin{verbatim}
% mtalk -modes UDP:TCP -ranking TCP
\end{verbatim}
will use UDP for gossip and TCP for all other communication. 
\end{description}


