\section{Identifiers}

\ensemble\ uses a variety of identifiers for a variety of different purposes.  Here
we summarize the important ones and describe what they are used for.  Their type
definitions can be found in the \sourcecode{type} directory in the file corresponding
to the name.  Look in \sourcetype{README} for an up to date listing of these files.
Most of the different identifiers support a similar interface for a variety of
operations.  Several of the identifiers are opaque, in the sense that the interface
hides the actual structure of the identifier.  All identifiers have a
\mlval{string\_of\_id} function defined which gives a human-readable description of
their contents.

\paragraph{Changes from \horus}
\begin{itemize}
\item
None of the identifiers are defined as fixed length sequences of bytes.  This is
something that remains to be done.
\item
\horus\ EID's have been split into endpoint and group identifiers.
\item
We have removed addressing information entirely from endpoint and group identifiers.
\item
Entity identifiers have been eliminated.  Connection, stack, and protocol identifiers
have been added to support a variety of features new to \ensemble.
\end{itemize}

\subsection{Endpoint Identifiers}
Endpoint identifiers are unique names for communication endpoints.  A single
process can create any number of local endpoint identifiers, each of which is
guaranteed to be unique (within some limits).  A process can have multiple
endpoints in a single group.  An endpoint can be a member of multiple groups.
However, the endpoints in a group must be distinct.

\subsection{Group Identifiers}
Group identifiers serve as unique names of communication groups.  They do not contain
addressing information.  The exception to this rule is that groups communicating via
Deering multicast choose a random multicast address by taking a hash of the group
address.  Processes can create any number of local group identifiers, each of which
is guaranteed to be unique (within some limits).

\subsection{View Identifiers}
View identifiers are unique identifiers of group views.  Whenever communication
protocols proceed through a view change, the resulting view is given a new view
identifier.  These are made unique by pairing the coordinator of the group with a
logical timestamp that is advanced whenever a view change happens.  Although two
partitions of a group may share the same time stamp, they will have different
coordinators.

\subsection{Connection Identifiers}
Connection identifiers are used to route messages to the precise destinations.  They
specify the exact destination endpoint or group, the view identifier, the protocol
stack to deliver to, the type of protocol being used to send the message, as well as
several other bits of information.  Typically, endpoint or group identifiers are used
to send messages to the correct processes and connection identifiers are used to
route messages to the exact destination of a message within a process.  Messages
usually contain a connection identifier as a ``header'' of the message but do not
contain endpoint identifiers or group identifiers, except as subfields of a
connection identifier.

\subsection{Protocol Identifiers}
There is a one-to-one relationship between the standard protocol stacks of \ensemble\
and protocol identifiers.  Applications select the protocol to use by specifying the
protocol id of that stack.  Having identifiers for protocols is convenient because
they can be passed around in messages and have equality comparisons made on them,
whereas the actual protocol stacks cannot.

\subsection{Mode identifiers}  
Each communication domain has a corresponding mode identifier used to specify that
domain.

\subsection{Stack Identifiers}
Stack identifiers are used to distinguish between the various domains that
a protocol stack may be receiving messages through.  Each of the various
kinds of ``channels'' that protocol stacks use has a separate identifier.
Currently, these are:
\begin{description}
\item
[Primary :] This is the primary communication channel for a protocol stack.
This is normally where most messages are received.
\item
[Bypass :] Messages sent via the optimized bypass protocols use this id. 
\item
[Merge :] Messages sent by group merge protocols use this id.
\item
[ProtoArb :] This is used for protocol-arbitration protocols (i.e.., protocols
that decide which protocol to use.
\item
[Authenticate :] This stack id is reserved for authentication protocols.
\end{description}
