\subsection{View state}

Several callbacks receive as an argument a pair of records with
information about the new view.  The information is split into two
parts, a \mlval{View.state} and a \mlval{View.local} record.  The
first contains information that is common to all the members in the
view, such as the \mlval{view} of the group.  The same record is
delivered to all members.  The second record contains information
local to the member that receives it.  These fields include the
\mlval{Endpt.id} of the member and its \mlval{rank} in the view.  It
also contains information that is derived from the \mlval{View.state}
record, such as \mlval{nmembers} with is merely the length of the
\mlval{view} field.

\begin{codebox}
(* VIEW.STATE: a record of information kept about views.
 * This value should be common to all members in a view.
 *)
type state = {
  (* Group information.
   *)
  version       : Version.id ;		(* version of Ensemble *)
  group		: Group.id ;		(* name of group *)
  proto_id	: Proto.id ;		(* id of protocol in use *)
  coord         : rank ;		(* initial coordinator *)
  ltime         : ltime ;		(* logical time of this view *)
  primary       : primary ;		(* primary partition? (only w/some protocols) *)
  groupd        : bool ;		(* using groupd server? *)
  xfer_view	: bool ;		(* is this an XFER view? *)
  key		: Security.key ;	(* keys in use *)
  prev_ids      : id list ;             (* identifiers for prev. views *)
  params        : Param.tl ;		(* parameters of protocols *)
  uptime        : Time.t ;		(* time this group started *)

  (* Per-member arrays.
   *)
  view 		: t ;			(* members in the view *)
  clients	: bool Arrayf.t ;	(* who are the clients in the group? *)
  address       : Addr.set Arrayf.t ;	(* addresses of members *)
  out_of_date   : ltime Arrayf.t	; (* who is out of date *)
  lwe           : Endpt.id Arrayf.t Arrayf.t ; (* for light-weight endpoints *)
  protos        : bool Arrayf.t  	(* who is using protos server? *)
}
\end{codebox}

\begin{codebox}
(* VIEW.LOCAL: information about a view that is particular to 
 * a member.
 *)
type local = {
  endpt	        : Endpt.id ;		(* endpoint id *)
  addr	        : Addr.set ;		(* my address *)
  rank 	        : rank ;		(* rank in the view *)  
  name		: string ;		(* my string name *)
  nmembers 	: nmembers ;		(* # members in view *)
  view_id 	: id ;			(* unique id of this view *)
  am_coord      : bool ;  		(* rank = vs.coord? *)
  falses        : bool Arrayf.t ;       (* all false: used to save space *)
  zeroes        : int Arrayf.t ;        (* all zero: used to save space *)
  loop          : rank Arrayf.t ;      	(* ranks in a loop, skipping me *)
  async         : (Group.id * Endpt.id) (* info for finding async *)
}  

(* LOCAL: create local record based view state and endpt.
 *)
val local : debug -> Endpt.id -> state -> local
\end{codebox}

Most of the fields are moderately self-explanatory.  If
\mlval{xfer\_view} is true, then this view is only for state transfer
and all members should take an \mlval{XferDone} action when the state
transfer is complete.  The view field is defined as \mlval{View.t},
which is:
\begin{codebox}
(* VIEW.T: an array of endpt id's.
 *)
type t = Endpt.id Arrayf.t
\end{codebox}
