%*************************************************************%
%
%    Ensemble, 1_42
%    Copyright 2003 Cornell University, Hebrew University
%           IBM Israel Science and Technology
%    All rights reserved.
%
%    See ensemble/doc/license.txt for further information.
%
%*************************************************************%
\begin{Layer}{ENCRYPT}

This layer encrypts application data for privacy.  Uses keys in the view state
record.  Authentication needs to be provided by the lower layers in the system.
The protocol headers are not encrypted.  This layer must reside above FIFO
layers for sending and receiving because it uses encryption contexts whereby
the encryption of a message is dependent on the previous messages from this
member.  These contexts are dropped at the end of a view.  A smarter protocol
would try to maintain them, as they improve the quality of the encryption.

\begin{Protocol}
Does chained encryption on the message payload in the \mlval{iov} field of
events.  Each member keeps track of the encryption state for all incoming and
outgoing point-to-point and multicast channels.  Messages marked
\mlval{Unreliable} are not encrypted (these should not be application
messages).
\end{Protocol}

\begin{Parameters}
\item None
\end{Parameters}

\begin{Properties}
\item
Guarantees (modulo encryption being broken) that only processes that know the
shared group key can read the contents of the application portion of data
messages.
\item
Requires FIFO ordering on point-to-point and multicast messages.
\end{Properties}

\begin{Sources}
\sourcesfile{encrypt.ml}
\end{Sources}

\begin{GenEvent}
\genevent{None}
\end{GenEvent}

\begin{Testing}
\item
see the VSYNC stack
\end{Testing}
\end{Layer}
