\begin{Layer}{INTER}

This protocol handles view changes that involve more than one
partition (see also INTRA).

\begin{Protocol}
Group merges are the more complicated part of the group membership
protocol.  However, we constrain the problem so that:
\begin{itemize}
\item
Groups cannot be both merging and accepting mergers at the same time.
This eliminates the potential for cycles in the ``merge-graph.''
\item
A view (i.e. view\_id) can only attempt to merge once, and only if no
failures have occured.  Each merge attempt is therefore uniquely
identified by the view\_id of the merging group.  Note also that by
requiring no failures to have occured for a merge to happen, this
prevents a member from being failed in one view and then reappearing
in the next view.  There has to be an intermediate view without the
failed member: this is a desirable property.
\end{itemize}
The merge protocol works as follows:
\begin{enumerate}
\item
The merging coordinator blocks its group,
\item
The merging coordinator sends a merge request to the remote group's
coordinator.
\item
The remote coordinator blocks its group,
\item
The remote coordinator installs a new view (with the mergers in it) and
sends the view to the merging coordinator (through a merge-granted
message).
\item
The merging coordinator installs the view in its group.
\end{enumerate}
If the merging coordinator times out on the merged coordinator then it
immediately installs a new view in its partition (without the other members
even finding out about the merge attempt).
\end{Protocol}

\begin{Parameters}
\item None
\end{Parameters}

\begin{Properties}
\item
When another partition is merging, a View message is also sent to the
coordinator of the merging group, which then forwards the message to
the rest of its group.
\item 
Requires that \Dn{Merge} events only be delivered by the original
coordinator of views (in which no failures have yet occured).
Otherwise, the partition should first form a new view and then attempt
the merge.
\item
\Dn{Merge} causes a \Dn{Merge} event to be delivered to the layer below.  This
will be replied with either an \Up{View}, \Up{MergeFailed}, or \Up{MergeDenied}
event, depending on the outcome of the merge attempt.
\item
\Up{MergeRequest}'s are only delivered at the coordinator.  And only if the
group is not currently blocking and only if the mergers list does not
contain members that are/were in this view or in previous merge requests in
this view.
\end{Properties}

\begin{Sources}
\sourcesfile{inter.ml}
\end{Sources}

\begin{GenEvent}
\genevent{\Dn{Merge}}
\genevent{\Dn{MergeDenied}}
\genevent{\Dn{Suspect}}
\end{GenEvent}

\begin{Testing}
\item
see the VSYNC stack
\end{Testing}
\end{Layer}

