%*************************************************************%
%
%    Ensemble, 2_00
%    Copyright 2004 Cornell University, Hebrew University
%           IBM Israel Science and Technology
%    All rights reserved.
%
%    See ensemble/doc/license.txt for further information.
%
%*************************************************************%
\begin{Layer}{LEAVE}

This protocol has two tasks.  (1) When a member really wants to leave a group,
the LEAVE protocol tells the other members to suspect this member.  (2) The
leave protocol garbage collects old protocol stacks by initiating a \Dn{Leave}
after getting an \Up{View} and then getting an \Up{Stable} where everything is
marked as being stable.

\begin{Protocol}
Both protocols are simple.

For leaving the group, a member broacasts a Leave message to the group
which causes the other members to deliver a \Dn{Suspect} event.  Note that
the other members will get the Leave message only after receiving all the
prior broadcast messages.  This member should probably stick around,
however, until these messages have stabilized.

Garbage collection is done by waiting until all broadcast message are
stable before delivering a local \Dn{Leave} event.
\end{Protocol}

\begin{Parameters}
\item
leave\_wait\_stable : whether or not to wait for the leave announcment to
become stable before leaving
\end{Parameters}

\begin{Properties}
\item \todo{}
\end{Properties}

\begin{Sources}
\sourcesfile{leave.ml}
\end{Sources}

\begin{GenEvent}
\genevent{\Dn{Leave}}
\end{GenEvent}

\begin{Testing}
\item see the VSYNC stack
\end{Testing}
\end{Layer}
