%*************************************************************%
%
%    Ensemble, 1_42
%    Copyright 2003 Cornell University, Hebrew University
%           IBM Israel Science and Technology
%    All rights reserved.
%
%    See ensemble/doc/license.txt for further information.
%
%*************************************************************%
\begin{Layer}{WINDOW} 

This layer implements window-based flow control based on stability information.
Multicast messages from each sender are sent only if the number of
unacknowledged messages from the sender is smaller than the window.

\begin{Protocol}
Whenever the number of unstable messages goes above the window, messages
are buffered without being sent.  On receipt of a stability update, the
number of unstable messages are recalculated and buffered messages are sent
as allowed by the window.
\end{Protocol}

\begin{Parameters}
\item window\_window : the window size in number of messages
\end{Parameters}

\begin{Properties}
\item
Requires stability information in the form of \Up{Stable} events.
\end{Properties}

\begin{Notes}
\item 
Future implementation should support dynamic window adjustment.
\item 
Performance with the WINDOW layer depends in part with the frequency of
stability updates.  The WINDOW flow control works the best when the
frequency is based on the number of unstable messages rather than on
periodic timeouts.
\item 
Alternative flow control layers include RATE and CREDIT.
\end{Notes}

\begin{Sources}
\sourcesfile{window.ml}
\end{Sources}

\emph{This layer and its documentation were written by Takako Hickey.}
\end{Layer}
